---
title: "Classification"
author: "Group 5 - Jonathan Che and David Green"
date: ""
output: 
  ptrain_document:
    fig_height: 3
    fig_width: 5
  html_document:
    fig_height: 3
    fig_width: 5
  word_document:
    fig_height: 3
    fig_width: 5
---


```{r, include=FALSE}
# Don't delete this chunk if you are using the mosaic package
# This loads the mosaic and dplyr packages
require(mosaic)
require(readr)
require(randomForest) # for random forests
```

```{r, include=FALSE}
# Some customization.  You can alter or delete as desired (if you know what you are doing).

# This changes the default colors in lattice plots.
trellis.par.set(theme=theme.mosaic())  

# knitr settings to control how R chunks work.
require(knitr)
opts_chunk$set(
  tidy=FALSE,     # display code as typed
  size="small"    # slightly smaller font for code
)
```

```{r, include=FALSE}
train <- read_csv("babynames4.csv")
test2012 <- read_csv("testdata2012.csv")
test2005 <- read_csv("testdata2005.csv")
test2000 <- read_csv("testdata2000.csv")
test1995 <- read_csv("testdata1995.csv")
test1900 <- read_csv("testdata1900.csv")
```

## Data Cleaning

```{r, warning=FALSE}
train <- data.frame(lapply(train[,1:3], function(x) as.factor(x)),
                  train[,4:6],
                  lapply(train[,7:8], function(x) as.factor(x)),
                  train[,9:10],
                  lapply(train[,11:13], function(x) as.logical(x)),
                  train[14:44])
train2011 <- train %>%
  filter(year == 2011)
train2011.rep <- train2011[rep(seq(dim(train2011)[1]), train2011$count), ]   # WE REPEAT THE DATA BY COUNT
train2011.rep.samp <- train2011.rep %>%
  sample_n(10000, replace=TRUE)   # Sample too large
train2011.final <- train2011.rep.samp %>%
  select(-year, -race, -name, -count, -rank)
rm(train2011.rep)
rm(train2011.rep.samp)
rm(train2011)
```

```{r}
clean_test_data <- function(df){
  df <- data.frame(df[,1],
                    lapply(df[,2], function(x) as.factor(x)),
                    df[,3:4],
                    lapply(df[,5:6], function(x) as.factor(x)),
                    df[,7:8],
                    lapply(df[,9:11], function(x) as.logical(x)),
                    df[12:42])
  df.rep <- df[rep(seq(dim(df)[1]), df$count_norm), ]   # WE REPEAT THE DATA BY COUNT
  df.final <- df.rep %>%
    select(-count_norm, -name, -count) %>%
    mutate(gender = ifelse(gender=="M", "MALE", "FEMALE")) %>%
    mutate(gender = as.factor(gender))
  levels(df.final$first_letter) <- levels(train2011.final$first_letter)
  levels(df.final$last_letter) <- levels(train2011.final$last_letter)
  return(df.final)
}
```

```{r}
test2012.final <- clean_test_data(test2012)
test2005.final <- clean_test_data(test2005)
test2000.final <- clean_test_data(test2000)
test1995.final <- clean_test_data(test1995)
test1900.final <- clean_test_data(test1900)
```


## Random Forest

> We train a model on the 2011 training data

```{r}
set.seed(40)
rf_2011=randomForest(gender ~., data=train2011.final,mtry=10,ntree=100,importance=T,proximity=T)
foo <- table(train2011.final$gender,predict(rf_2011,train2011.final)); foo
(foo[1,2]+foo[2,1])/sum(foo)   # Very low AER
varImpPlot(rf_2011)
```

> We build a function that allows us to analyze a given year

```{r}
analyze_year <- function(year){
  df <- get(paste("test", year, ".final", sep=""))
  table_name <- paste("table", year, sep="")
  error <- paste("error", year, sep="")
  errorm <- paste("error", year, "_m", sep="")
  errorf <- paste("error", year, "_f", sep="")
  
  assign(table_name, table(df$gender,predict(rf_2011, df)))
  table <- get(table_name)
  assign(error, (table[1,2]+table[2,1])/sum(table))
  assign(errorm, (table[2,1])/sum(table))
  assign(errorf, (table[1,2])/sum(table))
  
  return(c(table, get(error), get(errorm), get(errorf)))
}
```

> We test our model on the 2012 testing data

```{r}
info2012 <- analyze_year(2012)
rbind(info2012[c(1,3)], info2012[c(2,4)])
```

> We test our model on the 2005 testing data

```{r}
info2005 <- analyze_year(2005)
rbind(info2005[c(1,3)], info2005[c(2,4)])
```

> We test our model on the 2000 testing data

```{r}
info2000 <- analyze_year(2000)
rbind(info2000[c(1,3)], info2000[c(2,4)])
```

> We test our model on the 1995 testing data

```{r}
info1995 <- analyze_year(1995)
rbind(info1995[c(1,3)], info1995[c(2,4)])
```

> We test our model on the 1900 testing data

```{r}
info1900 <- analyze_year(1900)
rbind(info1900[c(1,3)], info1900[c(2,4)])
```

## Conclusions from Time Study

```{r}
years <- c(1900, 1995, 2000, 2005, 2012)
errors <- c(info1900[5], info1995[5], info2000[5], info2005[5], info2012[5])
errorsm <- c(info1900[6], info1995[6], info2000[6], info2005[6], info2012[6])
errorsf <- c(info1900[7], info1995[7], info2000[7], info2005[7], info2012[7])

df <- data.frame(years, errors, errorsm, errorsf)
ggplot(df, aes(x=years, y=errors)) +
  geom_point()
ggplot(df, aes(x=years, y=errorsm)) +
  geom_point()
ggplot(df, aes(x=years, y=errorsf)) +
  geom_point()
```

> We see that it gets more difficult over time, as expected.

> Very interestingly, it's only the female names that get more difficult over time. The male names remain similar.
