---
title: "Data Analysis"
author: "Group 4 - Jonathan Che and David Green"
date: ""
output: 
  pdf_document:
    fig_height: 3
    fig_width: 5
  html_document:
    fig_height: 3
    fig_width: 5
  word_document:
    fig_height: 3
    fig_width: 5
---


```{r, include=FALSE}
# Don't delete this chunk if you are using the mosaic package
# This loads the mosaic and dplyr packages
require(mosaic)
require(readr)
require(rpart) # for classification trees
require(class) # for kNN methods
require(randomForest) # for random forests
require(cluster)
```

```{r, include=FALSE}
# Some customization.  You can alter or delete as desired (if you know what you are doing).

# This changes the default colors in lattice plots.
trellis.par.set(theme=theme.mosaic())  

# knitr settings to control how R chunks work.
require(knitr)
opts_chunk$set(
  tidy=FALSE,     # display code as typed
  size="small"    # slightly smaller font for code
)
```

```{r, include=FALSE}
df <- read_csv("babynames4.csv")
```

## Data Cleaning

```{r, warning=FALSE}
df2 <- data.frame(lapply(df[,1:3], function(x) as.factor(x)),
                  df[,4:6],
                  lapply(df[,7:8], function(x) as.factor(x)),
                  df[,9:10],
                  lapply(df[,11:13], function(x) as.logical(x)),
                  df[14:44])

df3 <- df2 %>%
  select(-count, -rank, -name)

df2011 <- df3 %>%
  filter(year == "2011")
df2011a <- df2011 %>%   # Only numerical variables, no letter counts
  select(gender, name_length, vowel_consonant_prop, double_letter, double_vowel, 
         double_consonant, num_syllables, num_consonants, num_vowels,
         consonant_prop, vowel_prop)
df2011b <- df2011[,12:37]   # Only letter counts
```

## Visualizations

```{r}
# Insignificant differences unless otherwise noted
ggplot(df3, aes(x=gender, y=name_length)) +
  geom_boxplot()
ggplot(df3, aes(x=gender, y=vowel_consonant_prop)) +   # girls have more vowels/consonant
  geom_boxplot()
ggplot(df3, aes(x=gender, y=num_syllables)) +
  geom_boxplot()
ggplot(df3, aes(x=gender, y=name_length)) +
  geom_boxplot()
ggplot(df3, aes(x=gender, y=num_vowels)) +
  geom_boxplot()
ggplot(df3, aes(x=gender, y=num_consonants)) +
  geom_boxplot()

ggplot(df3, aes(x=name_length, y=num_syllables, color=race)) +
  geom_jitter()
```


## Clustering

```{r}
# http://finzi.psych.upenn.edu/library/ChemmineR/html/jarvisPatrick.html

# Clustering means that we don't want highly correlated variables, so we check correlations first
df2011.cluster1 <- df2011a %>%
  select(gender, vowel_prop, double_letter, name_length, num_syllables)
cor(df2011.cluster1[,-1], method="kendall")

# Heirarchical clustering on df2011.cluster1
dist1=dist(scale(df2011.cluster1[,-1])) #Statistical distances
hcward=hclust(dist1,method="ward.D") 
plot(hcward)   # 4 clusters emerge
hcward.sol=as.factor((cutree(hcward, k = 4)))
summary(hcward.sol)
foo <- data.frame(df2011, hcward.sol)

ggplot(foo, aes(x=hcward.sol, y=vowel_prop)) +
  geom_boxplot()
ggplot(foo, aes(x=hcward.sol, y=name_length)) +   # Cluster 2 has shorter names
  geom_boxplot()
ggplot(foo, aes(x=hcward.sol, y=num_syllables)) +
  geom_boxplot()
tally(gender~hcward.sol, data=foo)   # Cluster 4 is mostly male, 3 is balanced, 1&3 mostly female
tally(double_letter~hcward.sol, data=foo)   # Cluster 3 is the names with double letters
```

```{r}
# cor(df2011b, method="kendall")   # takes a long time to run, no significant correlations

# Heirarchical clustering on df2011.cluster1
dist2=dist(df2011b) #Statistical distances
hcward2=hclust(dist2, method="ward.D") 
plot(hcward2)   # 6ish clusters emerge
hcward.sol2=as.factor(cutree(hcward2, k = 6))
summary(hcward.sol2)
bar <- data.frame(df2011, hcward.sol2)

ggplot(bar, aes(x=hcward.sol2, y=num_consonants)) +   # Cluster 1 has more consonants, cluster 6 has fewer
  geom_boxplot()
ggplot(bar, aes(x=hcward.sol2, y=name_length)) +
  geom_boxplot()
ggplot(bar, aes(x=hcward.sol2, y=consonant_prop)) +
  geom_boxplot()

options(digits=1)
tally(a_count~hcward.sol2, data=bar, format="proportion")   # cluster 6 has more a
tally(l_count~hcward.sol2, data=bar, format="proportion")
tally(n_count~hcward.sol2, data=bar, format="proportion")
tally(e_count~hcward.sol2, data=bar, format="proportion")   # cluster 1 has more e


tally(gender~hcward.sol2, data=bar)   # nothing...
tally(double_letter~hcward.sol2, data=bar)   
```

## Simple Tree

```{r}
bf.control=rpart.control(minsplit=100,minbucket=20,xval=5)
bf.treeorig=rpart(gender ~.,data=df2011a,method="class",control=bf.control)
plot(bf.treeorig)
text(bf.treeorig,cex=.7)
printcp(bf.treeorig)
```

## Random Forest

```{r}
set.seed(40)
bf.rf=randomForest(gender ~., data=df2011a,mtry=3,ntree=100,importance=T,proximity=T)
bf.rf
table(df2011a$gender,predict(bf.rf,df2011a))

hist(treesize(bf.rf))
varImpPlot(bf.rf)
```

> With only a few variables, we achieve cross-validation error of 30%. Vowel/Consonant proportion is the most important variable here.

```{r}
set.seed(40)
bf.rf2=randomForest(gender ~., data=df2011,mtry=3,ntree=100,importance=T,proximity=T)
bf.rf2
table(df2011$gender,predict(bf.rf2,df2011))

hist(treesize(bf.rf2))
varImpPlot(bf.rf2)
```

> With all of the variables, the cross-validation error goes down to 12%. The first and last letter, as well as the amount of vowels matter a lot, but vowel/consonant proportion is still relatively important.

## NN Method

```{r}
bf.knn=knn(df2011a[,-1],df2011a[,-1],df2011[,"gender"],k=3,prob=T)
table(df2011a$gender,bf.knn)
```

> The nearest neighbor method gives similar results to the random forest (with the numerical variables only).


