---
title: "Data Analysis on Weighted Names"
author: "Group 5 - Jonathan Che and David Green"
date: ""
output: 
  pdf_document:
    fig_height: 3
    fig_width: 5
  html_document:
    fig_height: 3
    fig_width: 5
  word_document:
    fig_height: 3
    fig_width: 5
---


```{r, include=FALSE}
# Don't delete this chunk if you are using the mosaic package
# This loads the mosaic and dplyr packages
require(mosaic)
require(readr)
require(rpart) # for classification trees
require(class) # for kNN methods
require(randomForest) # for random forests
require(cluster)
```

```{r, include=FALSE}
# Some customization.  You can alter or delete as desired (if you know what you are doing).

# This changes the default colors in lattice plots.
trellis.par.set(theme=theme.mosaic())  

# knitr settings to control how R chunks work.
require(knitr)
opts_chunk$set(
  tidy=FALSE,     # display code as typed
  size="small"    # slightly smaller font for code
)
```

```{r, include=FALSE}
df <- read_csv("babynames4.csv")
```

## Data Cleaning

```{r, warning=FALSE}
df2 <- data.frame(lapply(df[,1:3], function(x) as.factor(x)),
                  df[,4:6],
                  lapply(df[,7:8], function(x) as.factor(x)),
                  df[,9:10],
                  lapply(df[,11:13], function(x) as.logical(x)),
                  df[14:44])
df2.rep <- df2[rep(seq(dim(df2)[1]), df2$count), ]   # WE REPEAT THE DATA BY COUNT
df2.rep.samp <- df2.rep %>%
  sample_n(10000, replace=TRUE)   # Sample too large

df3 <- df2.rep.samp %>%
  select(-count, -rank, -name)

df2011 <- df3 %>%
  filter(year == "2011")
df2011a <- df2011 %>%   # Only numerical variables, no letter counts
  select(gender, name_length, vowel_consonant_prop, double_letter, double_vowel, 
         double_consonant, num_syllables, num_consonants, num_vowels,
         consonant_prop, vowel_prop)
df2011b <- df2011[,12:37]   # Only letter counts
```

## Visualizations

```{r}
# Insignificant differences unless otherwise noted
ggplot (df3, aes(x=gender, y=name_length)) +
  geom_boxplot()
ggplot(df3, aes(x=gender, y=vowel_prop)) +  
  geom_boxplot() #girls have higher vowel/consonant ratio
ggplot(df3, aes(x=gender, y=num_syllables)) +
  geom_boxplot()
ggplot(df3, aes(x=gender, y=name_length)) +
  geom_boxplot()
ggplot(df3, aes(x=gender, y=num_vowels)) +
  geom_boxplot()
ggplot(df3, aes(x=gender, y=num_consonants)) +
  geom_boxplot()

ggplot(df3, aes(x=name_length, y=num_syllables, color=race)) +
  geom_jitter()
```

```{r}
tally(df2011$race, format="proportion")
tally(df2011$gender, format="proportion")
tally(df2011$gender ~ df2011$a_count, format="proportion")
tally(df2011$gender ~ df2011$b_count, format="proportion")
tally(df2011$gender ~ df2011$c_count, format="proportion")
tally(df2011$gender ~ df2011$d_count, format="proportion")
tally(df2011$gender ~ df2011$e_count, format="proportion")
tally(df2011$gender ~ df2011$f_count, format="proportion")
tally(df2011$gender ~ df2011$g_count, format="proportion")
tally(df2011$gender ~ df2011$h_count, format="proportion")
tally(df2011$gender ~ df2011$i_count, format="proportion")
tally(df2011$gender ~ df2011$j_count, format="proportion")
tally(df2011$gender ~ df2011$k_count, format="proportion")
tally(df2011$gender ~ df2011$l_count, format="proportion")
tally(df2011$gender ~ df2011$m_count, format="proportion")
tally(df2011$gender ~ df2011$n_count, format="proportion")
tally(df2011$gender ~ df2011$o_count, format="proportion")
tally(df2011$gender ~ df2011$p_count, format="proportion")
tally(df2011$gender ~ df2011$q_count, format="proportion")
tally(df2011$gender ~ df2011$r_count, format="proportion")
tally(df2011$gender ~ df2011$s_count, format="proportion")
tally(df2011$gender ~ df2011$t_count, format="proportion")
tally(df2011$gender ~ df2011$u_count, format="proportion")
tally(df2011$gender ~ df2011$v_count, format="proportion")
tally(df2011$gender ~ df2011$w_count, format="proportion")
tally(df2011$gender ~ df2011$x_count, format="proportion")
tally(df2011$gender ~ df2011$y_count, format="proportion")
tally(df2011$gender ~ df2011$z_count, format="proportion")
```

```{r}
#histogram(~vowel_consonant_prop | gender,data=df2011, width=.3)
#bwplot(~vowel_consonant_prop | gender, data=df2011, main="Vowel to Consonant Ratio in Girls and Boys")
ggplot(df3, aes(x=gender, y=vowel_consonant_prop,color=gender)) +
  geom_boxplot() +
  scale_color_manual(values=c("hotpink3","royalblue3")) +
  ggtitle("Vowel to Consonant Ratio in Girls and Boys")
```

```{r}
tally(gender~double_letter,data=df3)
tally(gender~double_consonant,data=df3)
ggplot(df3,aes(x=double_consonant,fill=gender)) +
  geom_bar(position="stack") +
  ggtitle("Amounts of Boy and Girl Names with and without Double-Consonants") #GIRLS HAVE MORE DOUBLE CONSONANTS (more significant than just double letters)
```

```{r}
tally(gender~num_vowels,data=df3)
ggplot(df3, aes(x=num_vowels,color=gender)) +
  geom_histogram(binwidth=1)
  # scale_color_manual(values=c("hotpink3","royalblue3")) +
  # ggtitle("Vowel to Consonant Ratio in Girls and Boys")
```

```{r}
ggplot(df3, aes(x=gender, y=vowel_prop,color=gender)) +
  geom_boxplot() +
  scale_color_manual(values=c("hotpink3","royalblue3")) +
  ggtitle("Vowel Proportion in Girls and Boys") #GIRLS HAVE HIGHER VOWEL PROP
```

```{r}
# create stacked proportional bar plots for every letter count with regard to gender (if you find-replace gender with race, you get the race plots)
temp <- df3
temp$a_count <- as.factor(temp$a_count) #MORE WOMEN
ggplot(temp,aes(x=gender,fill=a_count)) +
  geom_bar(position="fill") +
  ylab("proportion of that gender's names")
temp$b_count <- as.factor(temp$b_count)
ggplot(temp,aes(x=gender,fill=b_count)) +
  geom_bar(position="fill") +
  ylab("proportion of that gender's names")
temp$c_count <- as.factor(temp$c_count)
ggplot(temp,aes(x=gender,fill=c_count)) +
  geom_bar(position="fill") +
  ylab("proportion of that gender's names")
temp$d_count <- as.factor(temp$d_count) #MORE MEN
ggplot(temp,aes(x=gender,fill=d_count)) +
  geom_bar(position="fill") +
  ylab("proportion of that gender's names")
temp$e_count <- as.factor(temp$e_count)
ggplot(temp,aes(x=gender,fill=e_count)) +
  geom_bar(position="fill") +
  ylab("proportion of that gender's names")
temp$f_count <- as.factor(temp$f_count)
ggplot(temp,aes(x=gender,fill=f_count)) +
  geom_bar(position="fill") +
  ylab("proportion of that gender's names")
temp$g_count <- as.factor(temp$g_count)
ggplot(temp,aes(x=gender,fill=g_count)) +
  geom_bar(position="fill") +
  ylab("proportion of that gender's names")
temp$h_count <- as.factor(temp$h_count)
ggplot(temp,aes(x=gender,fill=h_count)) +
  geom_bar(position="fill") +
  ylab("proportion of that gender's names")
temp$i_count <- as.factor(temp$i_count) #MORE WOMEN
ggplot(temp,aes(x=gender,fill=i_count)) +
  geom_bar(position="fill") +
  ylab("proportion of that gender's names")
temp$j_count <- as.factor(temp$j_count) #MORE MEN
ggplot(temp,aes(x=gender,fill=j_count)) +
  geom_bar(position="fill") +
  ylab("proportion of that gender's names")
temp$k_count <- as.factor(temp$k_count)
ggplot(temp,aes(x=gender,fill=k_count)) +
  geom_bar(position="fill") +
  ylab("proportion of that gender's names")
temp$l_count <- as.factor(temp$l_count) #MORE WOMEN
ggplot(temp,aes(x=gender,fill=l_count)) +
  geom_bar(position="fill") +
  ylab("proportion of that gender's names")
temp$m_count <- as.factor(temp$m_count)
ggplot(temp,aes(x=gender,fill=m_count)) +
  geom_bar(position="fill") +
  ylab("proportion of that gender's names")
temp$n_count <- as.factor(temp$n_count) #MORE MEN
ggplot(temp,aes(x=gender,fill=n_count)) +
  geom_bar(position="fill") +
  ylab("proportion of that gender's names")
temp$o_count <- as.factor(temp$o_count) #MORE MEN
ggplot(temp,aes(x=gender,fill=o_count)) +
  geom_bar(position="fill") +
  ylab("proportion of that gender's names")
temp$p_count <- as.factor(temp$p_count)
ggplot(temp,aes(x=gender,fill=p_count)) +
  geom_bar(position="fill") +
  ylab("proportion of that gender's names")
temp$q_count <- as.factor(temp$q_count)
ggplot(temp,aes(x=gender,fill=q_count)) +
  geom_bar(position="fill") +
  ylab("proportion of that gender's names")
temp$r_count <- as.factor(temp$r_count)
ggplot(temp,aes(x=gender,fill=r_count)) +
  geom_bar(position="fill") +
  ylab("proportion of that gender's names")
temp$s_count <- as.factor(temp$s_count)
ggplot(temp,aes(x=gender,fill=s_count)) +
  geom_bar(position="fill") +
  ylab("proportion of that gender's names")
temp$t_count <- as.factor(temp$t_count)
ggplot(temp,aes(x=gender,fill=t_count)) +
  geom_bar(position="fill") +
  ylab("proportion of that gender's names")
temp$u_count <- as.factor(temp$u_count) #MORE MEN
ggplot(temp,aes(x=gender,fill=u_count)) +
  geom_bar(position="fill") +
  ylab("proportion of that gender's names")
temp$v_count <- as.factor(temp$v_count)
ggplot(temp,aes(x=gender,fill=v_count)) +
  geom_bar(position="fill") +
  ylab("proportion of that gender's names")
temp$w_count <- as.factor(temp$w_count)
ggplot(temp,aes(x=gender,fill=w_count)) +
  geom_bar(position="fill") +
  ylab("proportion of that gender's names")
temp$x_count <- as.factor(temp$x_count)
ggplot(temp,aes(x=gender,fill=x_count)) +
  geom_bar(position="fill") +
  ylab("proportion of that gender's names")
temp$y_count <- as.factor(temp$y_count) #MORE WOMEN
ggplot(temp,aes(x=gender,fill=y_count)) +
  geom_bar(position="fill") +
  ylab("proportion of that gender's names")
temp$z_count <- as.factor(temp$z_count)
ggplot(temp,aes(x=gender,fill=z_count)) +
  geom_bar(position="fill") +
  ylab("proportion of that gender's names")
```

## Clustering

```{r}
# http://finzi.psych.upenn.edu/library/ChemmineR/html/jarvisPatrick.html

# Clustering means that we don't want highly correlated variables, so we check correlations first
df2011.cluster1 <- df2011a %>%
  select(gender, vowel_prop, double_letter, name_length, num_syllables)
cor(df2011.cluster1[,-1], method="kendall")

# Heirarchical clustering on df2011.cluster1
dist1=dist(scale(df2011.cluster1[,-1])) #Statistical distances
hcward=hclust(dist1,method="ward.D") 
plot(hcward)   # 4 clusters emerge
hcward.sol=as.factor((cutree(hcward, k = 4)))
summary(hcward.sol)
foo <- data.frame(df2011, hcward.sol)

ggplot(foo, aes(x=hcward.sol, y=vowel_prop)) +
  geom_boxplot() #Cluster 1 has the highest vowel proportion
ggplot(foo, aes(x=hcward.sol, y=name_length)) +   
  geom_boxplot() #Clusters 1 thru 4 have progressively longer names
ggplot(foo, aes(x=hcward.sol, y=num_syllables)) +
  geom_boxplot()
tally(gender~hcward.sol, data=foo)   #Cluster 1 mostly women, 2 mostly men
tally(double_letter~hcward.sol, data=foo) #CLUSTER 3 MEANS DOUBLE LETTERS
```

```{r}
# cor(df2011b, method="kendall")   # takes a long time to run, no significant correlations

# Heirarchical clustering on df2011.cluster1
dist2=dist(df2011b) #Statistical distances
hcward2=hclust(dist2, method="ward.D") 
plot(hcward2)   # 6ish clusters emerge
hcward.sol2=as.factor(cutree(hcward2, k = 6))
summary(hcward.sol2)
bar <- data.frame(df2011, hcward.sol2)

ggplot(bar, aes(x=hcward.sol2, y=num_consonants)) +   
  geom_boxplot()
ggplot(bar, aes(x=hcward.sol2, y=name_length)) +
  geom_boxplot()
ggplot(bar, aes(x=hcward.sol2, y=consonant_prop)) +
  geom_boxplot()

options(digits=3)
tally(a_count~hcward.sol2, data=bar, format="proportion") #cluster 1 has more a
tally(l_count~hcward.sol2, data=bar, format="proportion")
tally(n_count~hcward.sol2, data=bar, format="proportion")
tally(r_count~hcward.sol2, data=bar, format="proportion")
tally(e_count~hcward.sol2, data=bar, format="proportion")   # cluster 2 has more e


tally(race~hcward.sol2, data=bar, format="proportion")   # nothing...
tally(double_letter~hcward.sol2, data=bar)   
```

## Simple Tree

```{r}
bf.control=rpart.control(minsplit=100,minbucket=20,xval=5)
bf.treeorig=rpart(gender ~.,data=df2011a,method="class",control=bf.control)
plot(bf.treeorig,margin=.1)
text(bf.treeorig,cex=.7)
printcp(bf.treeorig)
```

## Random Forest

```{r}
set.seed(40)
bf.rf=randomForest(gender ~., data=df2011a,mtry=3,ntree=100,importance=T,proximity=T)
bf.rf
table(df2011a$gender,predict(bf.rf,df2011a))

hist(treesize(bf.rf))
varImpPlot(bf.rf)
```

> With only a few variables, we achieve cross-validation error of 30%. Vowel/Consonant proportion is the most important variable here.

```{r}
set.seed(40)
bf.rf2=randomForest(gender ~., data=df2011,mtry=3,ntree=100,importance=T,proximity=T)
bf.rf2
table(df2011$gender,predict(bf.rf2,df2011))

hist(treesize(bf.rf2))
varImpPlot(bf.rf2)
```

> With all of the variables, the cross-validation error goes down to 12%. The first and last letter, as well as the amount of vowels matter a lot, but vowel/consonant proportion is still relatively important.

## NN Method

```{r}
bf.knn=knn(df2011a[,-1],df2011a[,-1],df2011[,"gender"],k=3,prob=T)
table(df2011a$gender,bf.knn)
```

> The nearest neighbor method gives similar results to the random forest (with the numerical variables only).

