---
title: "Progress Report"
author: "Group 4 - Jonathan Che and David Green"
date: "11/17/2016"
output: 
  pdf_document:
    fig_height: 3
    fig_width: 5
  html_document:
    fig_height: 3
    fig_width: 5
  word_document:
    fig_height: 3
    fig_width: 5
---


```{r, include=FALSE}
# Don't delete this chunk if you are using the mosaic package
# This loads the mosaic and dplyr packages
require(mosaic)
```

```{r, include=FALSE}
# Some customization.  You can alter or delete as desired (if you know what you are doing).

# This changes the default colors in lattice plots.
trellis.par.set(theme=theme.mosaic())  

# knitr settings to control how R chunks work.
require(knitr)
opts_chunk$set(
  tidy=FALSE,     # display code as typed
  size="small"    # slightly smaller font for code
)

```{r, include=FALSE}
# Load additional packages here.  Uncomment the line below to use Project MOSAIC data sets.
# require(mosaicData)   
```

## Work done so far

> Since the last update, we've wrapped up our preliminary feature extraction from the names. We've done a range of simple univariate and multivariate visualizations to help us get a grasp of what our data look like.

> We've also started fitting some simple classification models to classify names by gender or mother's race/ethnicity. So far, it seems that gender is a slightly easier classification to do. The classifications are going okay, but we can still do more experimentation.

## Work to do over Break

> We found an implementation of Jarvis-Patrick clustering at [http://finzi.psych.upenn.edu/library/ChemmineR/html/jarvisPatrick.html](http://finzi.psych.upenn.edu/library/ChemmineR/html/jarvisPatrick.html). We'll look into implementing it, as well as implementing other clustering algorithms, over break.

> We will also look into extracting more features from the names, to see if any of them can help with classification.

> Finally, we'll do things with the time information, such as:

1. Doing simple visualizations of changes in variables over time

2. Training models on a given year and testing them on the next year, then seeing what names these models get wrong to perhaps get some insight into changing name trends

